### 2022-09-11 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_networkary_classifiers@ae26545e508f95c05f6e91f39f9b29085a150e2a 🚀

\begin{enumerate}
    \item The prediction of species interactions is gaining momentum as
a way to circumvent limitations in data volume. Yet, ecological networks
are challenging to predict because they are typically small and sparse.
Dealing with extreme class imbalance is a challenge for most binary
classifiers, and there are currently no guidelines as to how predictive
models can be trained for this specific problem.%
    \item Using simple mathematical arguments and numerical experiments
in which a variety of classifiers (for supervised learning) are trained
on simulated networks, we develop a series of guidelines related to the
choice of measures to use for model selection, and the degree of
unbiasing to apply to the training dataset.%
    \item Neither classifier accuracy nor the ROC-AUC are informative
measures for the performance of interaction prediction. PR-AUC is a
fairer assessment of performance. In some cases, even standard measures
can lead to selecting a more biased classifier because the effect of
connectance is strong. The amount of correction to apply to the training
dataset depends on network connectance, on the measure to be optimized,
and only weakly on the classifier.%
    \item These results reveal that training machines to predict
networks is a challenging task, and that in virtually all cases, the
composition of the training set needs to be fine-tuned before performing
the actual training. We discuss these consequences in the context of the
low volume of data.%
\end{enumerate}

@@

Ecological networks are a backbone for key ecological and evolutionary
processes; yet enumerating all of the interactions between \(S\) species
is a daunting task, as it scales with \(S^2\), \emph{i.e.} the squared
species richness (Martinez, 1992). Recent contributions to the field of
ecological network prediction (Becker et al., 2022; Pichler et al.,
2020; Strydom et al., 2021) highlight that although interactions can be
predicted by adding ecologically relevant information (in the form of,
\emph{e.g.} traits), we do not have robust guidelines as to how the
predictive ability of these models should be evaluated, nor about how
the models should be trained. Here, by relying on simple derivations and
a series of simulations, we formulate a number of such guidelines,
specifically for the case of binary classifiers derived from thresholded
values. Specifically, we conduct an investigation of the models in terms
of their skill (ability to make the right prediction), bias (trends
towards systematically over-predicting one class), class imbalance (the
relative number of cases representing interactions), and show how these
effects interact. We conclude on the fact that models with the best
interaction-scale predictive score do not necessarily result in the most
accurate representation of the network.

The prediction of ecological interactions shares conceptual and
methodological issues with two fields in biology: species distribution
models (SDMs), and genomics. SDMs suffers from issues affecting
interactions prediction, namely low prevalence (due to sparsity of
observations/interactions) and data aggregation (due to bias in sampling
some locations/species). In previous work, Allouche et al. (2006)
suggested that \(\kappa\) was a better test of model performance than
the True Skill Statistic (TSS; which we refer to as Youden's
informedness); these conclusions were later criticized by Somodi et al.
(2017), who emphasized that informedness' is affected both by prevalence
and bias. Although this work offers recommendations about the comparison
of models, it doesn't establishes baselines or good practices for
training on imbalanced ecological data, or ways to remedy the imbalance.
Steen et al. (2021) show that, when applying spatial thinning (a process
that has no analogue in networks), the best approach to train ML-based
SDMs varies according to the balancing of the dataset, and the
evaluation measures used. This suggests that there is no single
``recipe'' that is guaranteed to give the best model. By contrast to
networks, SDMs have the advantage of being able to both thin datasets to
remove some of the sampling bias (\emph{e.g.} Inman et al., 2021), but
also to create pseudo-absences to inflate the number of supposed
negatives in the dataset (\emph{e.g.} Iturbide et al., 2015).

An immense body of research on machine learning application to life
sciences is focused on genomics (which has very specific challenges, see
a recent discussion by Whalen et al., 2021); this sub-field has
generated recommendations that do not necessarily match the current
best-practices for SDMs, and therefore hint at the importance of
domain-specific guidelines. Chicco \& Jurman (2020) suggest using
Matthews correlation coefficient (MCC) over \(F_1\), as a protection
against over-inflation of predicted results; Delgado \& Tibau (2019)
advocate against the use of Cohen's \(\kappa\), again in favor of MCC,
as the relative nature of \(\kappa\) means that a worse classifier can
be picked over a better one; similarly, Boughorbel et al. (2017)
recommend MCC over other measures of performance for imbalanced data, as
it has more desirable statistical properties. More recently, Chicco et
al. (2021) temper the apparent supremacy of the MCC, by suggesting it
should be replaced by Youden's informedness (also known as \(J\),
bookmaker's accuracy, and the True-Skill Statistic) when the imbalance
in the dataset may not be representative of the actual imbalance.

Species interaction networks are often under-sampled (Jordano, 2016a,
2016b), and this under-sampling is structured taxonomically (Beauchesne
et al., 2016), structurally (de Aguiar et al., 2019) and spatially
(Poisot, Bergeron, et al., 2021; Wood et al., 2015). As a consequence,
networks suffer from data deficiencies both within and between datasets.
This implies that the comparison of classifiers across space, when
undersampling varies locally (see \emph{e.g.} McLeod et al., 2021) is
non-trivial. Furthermore, the baseline value of classifiers performance
measures under various conditions of skill, bias, and prevalence, has to
be identified to allow researchers to evaluate whether their interaction
prediction model is indeed learning. Taken together, these
considerations highlight three specific issues for ecological networks.
First, what values of performance measures are indicative of a
classifier with no skill? This is particularly important as it can
evaluate whether low prevalence can lull us into a false sense of
predictive accuracy. Second, independently of the question of model
evaluation, is low prevalence an issue for \emph{training} or
\emph{testing}, and can we remedy it? Finally, because the low amount of
data on interaction makes a lot of imbalance correction methods (see
\emph{e.g.} Branco et al., 2015) hard to apply, which indicators can be
optimized by sacrificing least amount of positive interaction data?

It may sound counter-intuitive to care so deeply about how good a
classifier with no-skill is, as by definition, is has no skill. The
necessity of this exercise has its roots in the paradox of accuracy:
when the desired class (``two species interact'') is rare, a model that
gets less ecologically performant by only predicting the opposite class
(``these two species do not interact'') sees its accuracy increase;
because most of the guesses have ``these two species do not interact''
as a correct answer, a model that never predicts interactions would be
right an overwhelming majority of the time; it would also be utterly
useless. Herein lies the core challenge of predicting species
interactions: the extreme imbalance between classes makes the training
of predictive models difficult, and their validation even more so as we
do not reliably know which negatives are true. The connectance (the
proportion of realized interactions, usually the number of interactions
divided by the number of species pairs) of empirical networks is usually
well under 20\%, with larger networks having a lower connectance
(MacDonald et al., 2020), and therefore being increasingly difficult to
predict.

@@

Binary classifiers, which it to say, machine learning algorithms whose
answer is a categorical value, are usually assessed by measuring
properties of their confusion matrix, \emph{i.e.} the contingency table
reporting true/false positive/negative hits. A confusion matrix is laid
out as

@@

@@

A lot of binary classifiers are built by using a regressor (whose task
is to guess the value of the interaction, and can therefore return a
value considered to be a pseudo-probability); in this case, the optimal
value below which predictions are assumed to be negative (\emph{i.e.}
the interaction does not exist) can be determined by picking a threshold
maximizing some value on the ROC or the PR curve. The area under these
curves (ROC-AUC and PR-AUC henceforth) give ideas on the overall
goodness of the classifier, and the ideal threshold is the point on
these curves that minimizes the tradeoff represented in these curves.
Saito \& Rehmsmeier (2015) established that the ROC-AUC is biased
towards over-estimating performance for imbalanced data; on the
contrary, the PR-AUC is able to identify classifiers that are less able
to detect positive interactions correctly, with the additional advantage
of having a baseline value equal to prevalence. Therefore, it is
important to assess whether these two measures return different results
when applied to ecological network prediction. The ROC curve is defined
by the false positive rate on the \(x\) axis, and the true positive rate
on the \(y\) axis, and the PR curve is defined by the true positive rate
on the \(x\) axis, and the positive predictive value on the \(y\) axis.
By comparison with the previous paragraph, it is obvious that \(F_1\)
and MCC have ties to the PR curve (being close to the expected PR-AUC),
and that informedness has ties to the ROC curve (whereby the threshold
maximizing informedness is also the point of maximal inflection on the
ROC curve). One important difference between ROC and PR is that the
later does not prominently account for the size of the true negative
compartments: in short, it is more sensitive to the correct positive
predictions. In a context of strong imbalance, PR-AUC is therefore a
more stringent test of model performance.

@@

In this section, we will assume a network of connectance \(\rho\),
\emph{i.e.} having \(\rho S^2\) interactions (where \(S\) is the species
richness), and \((1-\rho) S^2\) non-interactions. Therefore, the vector
describing the \emph{true} state of the network (assumed to be an
unweighted, directed network) is a column vector
\(\mathbf{o}^T = [\rho, (1-\rho)]\) (we can safely drop the \(S^2\)
terms, as we will work on the confusion matrix, which ends up expressing
\emph{relative} values). We will apply skill and bias to this matrix,
and measure how a selection of performance metrics respond to changes in
these values, in order to assess their suitability for model evaluation.

@@

The training sample is composed of 50\% of the \(10^4\) possible entries
in the network, \emph{i.e.} \(n=5000\). Out of these interactions, we
pick a proportion \(\nu\) (the training set balance) to be positive, so
that the training set has \(\nu n\) interactions, and \((1-\nu) n\)
non-interactions. We vary \(\nu\) uniformly in \(]0,1[\). This allows to
evaluate how the measures of binary classification performance respond
to artificially rebalanced dataset for a given network connectance. The
rest of the dataset is used as a testing set, on which all further
measures are calculated. Note that although the training set is balanced
arbitrarily, the testing set is assembled so that it has the exact
connectance of the entire network; this ensures that the model is
evaluated under the class imbalance where the predictions will be made,
which represents a more meaningful evaluation. Note also that although
the simulated networks are bipartite, the algorithms have no
``knowledge'' of the network structure, and simply look at pairs of
species; therefore, the approach outlined here would also work for
unipartite networks.

The dataset used for numerical experiments is composed of a grid of 35
values of connectance (from 0.011 to 0.5) and 35 values of \(\nu\) (from
0.02 to 0.98); for each pair of values, 500 networks are generated and
predicted. For each network, we train four machines: a trait-based k-NN
(\emph{e.g.} Desjardins-Proulx et al., 2017), a regression tree, a
regression random forest, and a boosted regression tree. Following
results from Pichler et al. (2020), linear models have not been
considered (in any way, the relationship in the simulated networks is
non-linear). The point of these numerical experiments is \emph{not} to
recommend the best model (this is likely problem-specific), but to
highlight a series of recommendations that would work for supervised
learning tasks. All models were taken from the \texttt{MLJ.jl} package
(Blaom et al., 2020; Blaom \& Vollmer, 2020) in Julia 1.7 (Bezanson et
al., 2017). All machines use the default parameterization; this is an
obvious deviation from best practices, as the hyperparameters of any
machine require training before its application on a real dataset. As we
use 612500 such datasets, this would require over 2 millions unique
instances of tweaking the hyperparameters, which is prohibitive from a
computing time point of view. An important thing to keep in mind is that
the problem we simulate has been designed to be simple to solve: we
expect all machines with sensible default parameters to fare well ---
the results presented in the later sections show that this assumption is
warranted, and we further checked that the models do not overfit by
ensuring that there is never more than 5\% of difference between the
accuracy on the training and testing sets. All machines return a
quantitative prediction, usually (but not necessarily) in \([0,1]\),
which is proportional (but not necessarily linearly) to the probability
of an interaction between \(i\) and \(j\). The ROC-AUC and PR-AUC (and
therefore the thresholds) can be measured by integrating over the domain
of the values return by each machine, but in order to make the
average-based ensemble model more meaningful, all predictions are
expressed in \([0,1]\).

@@

It is noteworthy that the ensemble model was systematically better than
the component models. We do not expect that ensembles will \emph{always}
be better than single models. Networks with different structures than
the one we simulated here may respond in different ways, especially if
the rules are fuzzier than the simple rule we used here. In a recent
multi-model comparison involving supervised and unsupervised learning,
Becker et al. (2022) found that the ensemble was \emph{not} the best
model, and was specifically under-performing compared to models using
biological traits. This may be because the dataset of Becker et al.
(2022) was known to be under-sampled, and so the network itself
contained less information than the network and species traits. There is
no general conclusion to draw from either these results or ours, besides
reinforcing the need to be pragmatic about which models should be
included in the ensemble, and whether to use an ensemble at all. In a
sense, the surprising performance of the ensemble model should form the
basis of the first broad recommendation: optimal training set balance
and its interaction with connectance and the specific binary classifier
used is, in a sense, an hyperparameter that should be assessed. The
distribution of results in fig.~\ref{fig:optimbias} and
fig.~\ref{fig:optimvalue} show that there are variations around the
trend, and multiple models should probably be trained on their
``optimal'' training/testing set, as opposed to the same ones.

The results presented here highlight an interesting paradox: although
the k-NN model was ultimately able to get a correct estimate of network
structure (see tbl.~\ref{tbl:comparison} and fig.~\ref{fig:ecovalid}),
it ultimately remains a poor classifier, as evidenced by its low PR-AUC.
This suggests that the goal of predicting \emph{interactions} and
predicting \emph{networks} may not always be solvable in the same way --
of course a perfect classifier of interactions would make a perfect
network prediction; indeed, the best scoring predictor of interactions
(the ensemble model) had the best prediction of network structure. The
tasks of predicting networks structure and of predicting interactions
within networks are essentially two different ones. For some
applications (\emph{.e.g.} comparison of network structure across
gradients), one may care more about a robust estimate of the structure,
at the cost at putting some interactions at the wrong place. For other
applications (\emph{e.g.} identifying pairs of interacting species), one
may conversely care more about getting as many pairs right, even though
the mistakes accumulate in the form of a slightly worse estimate of
network structure. How these two approaches can be reconciled is
something to evaluate on a case-by-case basis, especially since there is
no guarantee that an esemble model will always be the most precise one.
Despite this apparent tension at the heart of the predictive exercise,
we can use the results presented here to suggest a number of guidelines.

First, because we have more trust in reported interactions than in
reported absences of interactions (which are overwhelmingly
\emph{pseudo}-absences), we can draw on previous literature to recommend
informedness as a measure to decide on a threshold for binary
classification (Chicco et al., 2021); this being said, because
informedness is insensitive to bias (although it is a linear measure of
skill), the overall model performance is better evaluated through the
use of MCC (figs.~\ref{fig:optimbias}, \ref{fig:optimvalue}). Because
\(F_1\) is monotonously sensitive to classifier bias fig.~\ref{fig:bias}
and network connectance fig.~\ref{fig:connectance}, MCC should be
prefered as a measure of model evaluation and comparison. When dealing
with multiple models, we therefore suggest to find the optimal threshold
using informedness, and to pick the best model using MCC (assuming one
does not want to use an ensemble model).

@@

One key element for real-life data that can make the prediction exercise
more tractable is that some interactions can safely be assumed to be
impossible; indeed, a lot of networks admit a stochastic block model as
a good approximation (\emph{e.g.} Xie et al., 2017). In ecological
networks, this can be due to spatial constrains (Valdovinos, 2019), or
to the long-standing knowledge that some links are ``forbidden'' due to
traits (Olesen et al., 2011) or abundances (Canard et al., 2014). The
matching rules (Olito \& Fox, 2015; Strona \& Veech, 2017) can be
incorporated in the model either by adding compatibility traits, or by
\emph{only} training the model on pairs of species that are not likely
to be forbidden links; having this information would allow to assemble
training/testing sets that have true negatives, and in this situation,
it may be possible to use the more usual 70/30 split. Besides forbidden
links, a real-life case that may arise is multi-interaction or
multi-layer networks (Pilosof et al., 2017). These can be studied using
the same general approach outlined here, either by assuming that pairs
of species can interact in more than one way (wherein one would train a
model for each type of interaction, based on the relevant predictors),
or by assuming that pairs of species can only have one type of
interaction (wherein this becomes a multi-label classification problem).

@@

>>>>

\begin{enumerate}
    \item The prediction of species interactions is gaining momentum as
a way to circumvent limitations in data volume. Yet, ecological networks
are challenging to predict because they are typically small and sparse.
Dealing with extreme class imbalance is a challenge for most binary
classifiers, and there are currently no guidelines as to how predictive
models can be trained for this specific problem.%
    \item Using simple mathematical arguments and numerical experiments
in which a variety of classifiers (for supervised learning) are trained
on simulated networks, we develop a series of guidelines related to the
choice of measures to use for model selection, and the ways to assemble
the training dataset.%
    \item Neither classifier accuracy nor the area under the receiver
operating characteristic curve (ROC-AUC) are informative measures for
the performance of interaction prediction. The area under the
precision-recall curve (PR-AUC) is a fairer assessment of performance.
In some cases, even standard measures can lead to selecting a more
biased classifier because the effect of connectance is strong. The
amount of correction to apply to the training dataset depends on network
connectance, on the measure to be optimized, and only weakly on the
classifier.%
    \item These results reveal that training machines to predict
networks is a challenging task, and that in virtually all cases, the
composition of the training set needs to be fine-tuned before performing
the actual training. We discuss these consequences in the context of the
low volume of data.%
\end{enumerate}

@@

Species interactions, forming ecological networks, are a backbone for
key ecological and evolutionary processes; yet enumerating all of the
interactions between \(S\) species is a daunting task, as it scales with
\(S^2\), \emph{i.e.} the squared species richness (Martinez, 1992).
Recent contributions to the field of ecological network prediction
(Becker et al., 2022; Pichler et al., 2020; Strydom et al., 2021)
highlight that although interactions can be predicted by adding
ecologically relevant information (in the form of, \emph{e.g.} traits),
we do not have robust guidelines as to how the predictive ability of
models recommending species interactions should be evaluated, nor about
how these models should be trained. Here, by relying on simple
derivations and a series of simulations, we formulate a number of such
guidelines, specifically for the case of binary classifiers derived from
thresholded values. Specifically, we conduct an investigation of the
models in terms of their skill (ability to make the right prediction),
bias (trends towards systematically over-predicting one class), class
imbalance (the relative number of cases representing interactions), and
show how these effects interact. We conclude on the fact that models
with the best interaction-scale predictive score do not necessarily
result in the most accurate representation of the true network.

The prediction of ecological interactions shares conceptual and
methodological issues with two fields in biology: species distribution
modelling (SDMs), and genomics. SDMs suffers from issues affecting
interactions prediction, namely low prevalence (due to sparsity of
observations/interactions) and data aggregation (due to bias in sampling
some locations/species). An important challenge lies in the fact that
the best measure to quantify the performance of a model is not
necessarilly a point of consensus (these methods, their interpretation,
and the way they are measured, are covered in depth in the next
section). In previous work, Allouche et al. (2006) suggested that
Cohen's \(\kappa\) agreement score (\(\kappa\) thereafter) was a better
test of model performance than the True Skill Statistic (TSS; which we
refer to as Youden's informedness thereafter); these conclusions were
later criticized by Somodi et al. (2017), who emphasized that
informedness is affected both by prevalence and bias. Although this work
offers recommendations about the comparison of models, it doesn't
establishes baselines or good practices for training on imbalanced
ecological data, or ways to remedy the imbalance. Steen et al. (2021)
show that, when applying spatial thinning (artificially re-balancing
observation data in space to avoid artifacts due to auto-correlation),
the best approach to train ML-based SDMs varies according to the
balancing of the dataset, and the evaluation measures used; there is no
single ``recipe'' that is guaranteed to give the best model. By contrast
to networks, SDMs have the advantage of being able to both thin datasets
to remove some of the sampling bias (\emph{e.g.} Inman et al., 2021),
but also to create pseudo-absences to inflate the number of supposed
negatives in the dataset (\emph{e.g.} Iturbide et al., 2015). These
powerful ways to remove data bias often have no analogue in networks,
removing one potential tool from our methodological toolkit, and making
the task of network prediction through classification potentially more
demanding, and more prone to underlying data biases.

An immense body of research on machine learning application to life
sciences is focused on genomics (which has very specific challenges, see
a recent discussion by Whalen et al., 2021); this sub-field has
generated recommendations that do not necessarily match the current
best-practices for SDMs, and therefore hint at the importance of
domain-specific guidelines. Chicco \& Jurman (2020) suggest using
Matthews correlation coefficient (MCC) over \(F_1\), as a protection
against over-inflation of predicted results; Delgado \& Tibau (2019)
advocate against the use of Cohen's \(\kappa\), again in favor of MCC,
as the relative nature of \(\kappa\) means that a worse classifier can
be picked over a better one; similarly, Boughorbel et al. (2017)
recommend MCC over other measures of performance for imbalanced data, as
it has more desirable statistical properties. More recently, Chicco et
al. (2021) temper the apparent supremacy of the MCC, by suggesting it
should be replaced by Youden's informedness (also known as \(J\),
bookmaker's accuracy, and the True-Skill Statistic) when the imbalance
in the dataset may not be representative of the actual imbalance. In a
way, the measures themselves need not be a strong focus for network
prediction, as they are routinely used in other field; the
discipline-specific question we seek to address is: `which metric should
be employed when predicting networks, and how to optimize it?'

Species interaction networks are often under-sampled (Jordano, 2016a,
2016b), and this under-sampling is structured taxonomically (Beauchesne
et al., 2016), structurally (de Aguiar et al., 2019) and spatially
(Poisot, Bergeron, et al., 2021; Wood et al., 2015). As a consequence,
networks suffer from data deficiencies both within and between datasets.
This implies that the comparison of classifiers across space, when
undersampling varies locally (see \emph{e.g.} McLeod et al., 2021) is
non-trivial. Furthermore, the baseline value of classifiers performance
measures under various conditions of skill, bias, and prevalence, has to
be identified to allow researchers to evaluate whether their interaction
prediction model is indeed learning. Taken together, these
considerations highlight three specific issues for ecological networks.
First, what values of performance measures are indicative of a
classifier with no skill? This is particularly important as it can
evaluate whether low prevalence can lull us into a false sense of
predictive accuracy. Second, independently of the question of model
evaluation, is low prevalence an issue for \emph{training} or
\emph{testing}, and can we remedy it? Finally, because the low amount of
data on interaction makes a lot of imbalance correction methods (see
\emph{e.g.} Branco et al., 2015) hard to apply, which measures of model
performance can be optimized by sacrificing least amount of positive
interaction data?

A preliminary question is to examin the baseline performance of these
measures, \emph{i.e.} the values they would take on hypothetical
networks based on a classifier that has no-skill. It may sound
counter-intuitive to care so deeply about how good a classifier with
no-skill is, as by definition, is has no skill. The necessity of this
exercise has its roots in the paradox of accuracy: when the desired
class (``two species interact'') is rare, a model that gets less
ecologically performant by only predicting the opposite class (``these
two species do not interact'') sees its accuracy increase; because most
of the guesses have ``these two species do not interact'' as a correct
answer, a model that never predicts interactions would be right an
overwhelming majority of the time; it would also be utterly useless.
Herein lies the core challenge of predicting species interactions: the
extreme imbalance between classes makes the training of predictive
models difficult, and their validation even more so as we do not
reliably know which negatives are true. The connectance (the proportion
of realized interactions, usually the number of interactions divided by
the number of species pairs) of empirical networks is usually well under
20\%, with larger networks having a lower connectance (MacDonald et al.,
2020), and therefore being increasingly difficult to predict.

@@

Binary classifiers, which it to say, machine learning algorithms whose
answer is a binary value, are usually assessed by measuring properties
of their confusion matrix, \emph{i.e.} the contingency table reporting
true/false positive/negative hits. A confusion matrix is laid out as

@@

A lot of binary classifiers are built by using a regressor (whose task
is to guess the value of the interaction, and can therefore return a
value considered to be a pseudo-probability); in this case, the optimal
value below which predictions are assumed to be negative (\emph{i.e.}
the interaction does not exist) can be determined by picking a threshold
maximizing some value on the ROC or the PR curve. The area under these
curves (ROC-AUC and PR-AUC henceforth) give ideas on the overall
goodness of the classifier, and the ideal threshold is the point on
these curves that minimizes the tradeoff represented in these curves.
Saito \& Rehmsmeier (2015) established that the ROC-AUC is biased
towards over-estimating performance for imbalanced data; on the
contrary, the PR-AUC is able to identify classifiers that are less able
to detect positive interactions correctly, with the additional advantage
of having a baseline value equal to prevalence. Therefore, it is
important to assess whether these two measures return different results
when applied to ecological network prediction. The ROC curve is defined
by the false positive rate on the \(x\) axis, and the true positive rate
on the \(y\) axis, and the PR curve is defined by the true positive rate
on the \(x\) axis, and the positive predictive value on the \(y\) axis.

@@

One noteworthy fact is that \(F_1\) and MCC have ties to the PR curve
(being close to the expected PR-AUC), and that informedness has ties to
the ROC curve (whereby the threshold maximizing informedness is also the
point of maximal inflection on the ROC curve). One important difference
between ROC and PR is that the later does not prominently account for
the size of the true negative compartments: in short, it is more
sensitive to the correct positive predictions. In a context of strong
imbalance, PR-AUC is therefore a more stringent test of model
performance.

@@

In this section, we will assume a network with connectance equal to a
scalar \(\rho\), \emph{i.e.} having \(\rho S^2\) interactions (where
\(S\) is the species richness), and \((1-\rho) S^2\) non-interactions.
Therefore, the vector describing the \emph{true} state of the network
(assumed to be an unweighted, directed network) is a column vector
\(\mathbf{o}^T = [\rho, (1-\rho)]\) (we can safely drop the \(S^2\)
terms, as we will work on the confusion matrix, which ends up expressing
\emph{relative} values). We will apply skill and bias to this matrix,
and measure how a selection of performance metrics respond to changes in
these values, in order to assess their suitability for model evaluation.

@@

The training sample is composed of a random pick of up to 50\% of the
\(10^4\) possible entries in the network, \emph{i.e.} \(n=5000\). Out of
these interactions, we pick a proportion \(\nu\) (the training set
balance) to be positive, so that the training set has \(\nu n\)
interactions, and \((1-\nu) n\) non-interactions. We vary \(\nu\)
uniformly in \(]0,1[\). This allows to evaluate how the measures of
binary classification performance respond to artificially rebalanced
dataset for a given network connectance. The rest of the dataset is used
as a testing set, on which all further measures are calculated. Note
that although the training set is balanced arbitrarily, the testing set
is assembled so that it has the exact connectance of the entire network;
this ensures that the model is evaluated under the class imbalance where
the predictions will be made, which represents a more meaningful
evaluation. Furthermore, to avoid artifacts due to different sizes of
the training and testing set within a single network, the number of
entries in both sets are equal. Note also that although the simulated
networks are bipartite, the algorithms have no ``knowledge'' of the
network structure, and simply look at pairs of species; therefore, the
approach outlined here would also work for unipartite networks.

The dataset used for numerical experiments is composed of a grid of 35
values of connectance (from 0.011 to 0.5) and 35 values of \(\nu\) (from
0.02 to 0.98); for each pair of values, 500 networks are generated and
predicted. For each network, we train four machines: a trait-based k-NN
(\emph{e.g.} Desjardins-Proulx et al., 2017), a regression tree, a
regression random forest, and a boosted regression tree; the later three
methods are turned into classifiers using thresholding, which oftentimes
provides better results than classification when faced with class
imbalance (Hong et al., 2016). Following results from Pichler et al.
(2020), linear models have not been considered (in any way, the
relationship in the simulated networks is non-linear). The point of
these numerical experiments is \emph{not} to recommend the best model
(this is likely problem-specific), but to highlight a series of
recommendations that would work for supervised learning tasks. All
models were taken from the \texttt{MLJ.jl} package (Blaom et al., 2020;
Blaom \& Vollmer, 2020) in Julia 1.7 (Bezanson et al., 2017). All
machines use the default parameterization; this is an obvious deviation
from best practices, as the hyperparameters of any machine require
training before its application on a real dataset. As we use 612500 such
datasets, this would require over 2 millions unique instances of
tweaking the hyperparameters, which is prohibitive from a computing time
point of view. An important thing to keep in mind is that the problem we
simulate has been designed to be simple to solve: we expect all machines
with sensible default parameters to fare well --- the results presented
in the later sections show that this assumption is warranted, and we
further checked that the models do not overfit by ensuring that there is
never more than 5\% of difference between the accuracy on the training
and testing sets. All machines return a quantitative prediction, usually
(but not necessarily) in \([0,1]\), which is proportional (but not
necessarily linearly) to the probability of an interaction between \(i\)
and \(j\). The ROC-AUC and PR-AUC (and therefore the thresholds) can be
measured by integrating over the domain of the values return by each
machine, but in order to make the average-based ensemble model more
meaningful, all predictions are expressed in \([0,1]\).

@@

It is noteworthy that the ensemble model was systematically better than
the component models. We do not expect that ensembles will \emph{always}
be better than single models. Networks with different structures than
the one we simulated here may respond in different ways, especially if
the rules are fuzzier than the simple rule we used here. In a recent
multi-model comparison involving supervised and unsupervised learning,
Becker et al. (2022) found that the ensemble was \emph{not} the best
model, and was specifically under-performing compared to models using
biological traits. This may be because the dataset of Becker et al.
(2022) was known to be under-sampled, and so the network alone contained
less information than the combination of the network and species traits.
There is no general conclusion to draw from either these results or
ours, besides reinforcing the need to be pragmatic about which models
should be included in the ensemble, and whether to use an ensemble at
all. In a sense, the surprising performance of the ensemble model should
form the basis of the first broad recommendation: optimal training set
balance and its interaction with connectance and the specific binary
classifier used is, in a sense, an hyperparameter that should be
assessed following the approach outlined in this manuscript. The
distribution of results in fig.~\ref{fig:optimbias} and
fig.~\ref{fig:optimvalue} show that there are variations around the
trend, and multiple models should probably be trained on their
``optimal'' training/testing set, as opposed to the same ones.

The results presented here highlight an interesting paradox: although
the k-NN model was ultimately able to get a correct estimate of network
structure (see tbl.~\ref{tbl:comparison} and fig.~\ref{fig:ecovalid}),
it ultimately remains a poor classifier, as evidenced by its low PR-AUC.
This suggests that the goal of predicting \emph{interactions} and
predicting \emph{networks} may not always be solvable in the same way --
of course a perfect classifier of interactions would make a perfect
network prediction; indeed, the best scoring predictor of interactions
(the ensemble model) had the best prediction of network structure. The
tasks of predicting networks structure and of predicting interactions
within networks are essentially two different ones. For some
applications (\emph{e.g.} comparison of network structure across
gradients), one may care more about a robust estimate of the structure,
at the cost at putting some interactions at the wrong place. For other
applications (\emph{e.g.} identifying pairs of interacting species), one
may conversely care more about getting as many pairs right, even though
the mistakes accumulate in the form of a slightly worse estimate of
network structure. How these two approaches can be reconciled is
something to evaluate on a case-by-case basis, especially since there is
no guarantee that an ensemble model will always be the most precise one.
Despite this apparent tension at the heart of the predictive exercise,
we can use the results presented here to suggest a number of guidelines.

First, because we have more trust in reported interactions than in
reported absences of interactions (which are overwhelmingly
\emph{pseudo}-absences), we can draw on previous literature to recommend
informedness as a measure to decide on a threshold for binary
classification (Chicco et al., 2021); this being said, because
informedness is insensitive to bias (although it is a linear measure of
skill), the overall model performance is better evaluated through the
use of MCC (figs.~\ref{fig:optimbias}, \ref{fig:optimvalue}). Because
\(F_1\) is monotonously sensitive to classifier bias
(fig.~\ref{fig:bias}) and network connectance
(fig.~\ref{fig:connectance}), MCC should be prefered as a measure of
model evaluation and comparison. When dealing with multiple models, we
therefore suggest to find the optimal threshold using informedness, and
to pick the best model using MCC (assuming one does not want to use an
ensemble model).

@@

One key element for real-life data that can make the prediction exercise
more tractable is that some interactions can safely be assumed to be
impossible; indeed, a lot of networks can be reasonably well described
using a stochastic block model (\emph{e.g.} Xie et al., 2017). In
ecological networks, this can be due to spatial constraints (Valdovinos,
2019), or to the long-standing knowledge that some links are
``forbidden'' due to traits (Olesen et al., 2011) or abundances (Canard
et al., 2014). The matching rules (Olito \& Fox, 2015; Strona \& Veech,
2017) can be incorporated in the model either by adding compatibility
traits, or by \emph{only} training the model on pairs of species that
are not likely to be forbidden links. Knowledge of true negative
interactions could be propagated in training/testing sets that have true
negatives, and in this situation, it may be possible to use the more
usual 70/30 split for training/testing folds as the need to protect
against potential unbalance is lowered. Besides forbidden links, a
real-life case that may arise is multi-interaction or multi-layer
networks (Pilosof et al., 2017). These can be studied using the same
general approach outlined here, either by assuming that pairs of species
can interact in more than one way (wherein one would train a model for
each type of interaction, based on the relevant predictors), or by
assuming that pairs of species can only have one type of interaction
(wherein this becomes a multi-label classification problem).

@@

\leavevmode\hypertarget{ref-Hong2016DeaCla}{}%
Hong, C., Ghosh, R., \& Srinivasan, S. (2016). \emph{Dealing with Class
Imbalance using Thresholding}.
\url{https://doi.org/10.48550/arXiv.1607.02705}
